Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-16 WholePlotTrajectories.Rmd
Discussion

Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)). 
*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*

@@

For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig6)). 
Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.

@@

The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2010).
Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).

The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth with a sharp increase until 15 years after disturbance.
Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
Along with the IDH, disturbance then enhanced the growth and survival of previously infrequent species, reorganizing the typical high dominance structure of hyperdiverse mature forests.
This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
In line with the IDH, the post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].

@@

The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
All plots, though, followed the same unimodal trajectory which confirmed both the resilience of forests taxonomic and functional composition and the convergence of communities towards initial states [@Hubbell1999;@Molino2001;@Baraloto2012a].
Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.

@@

After disturbance, communities functional characteristics followed a similar but delayed trajectory as this of taxonomic characteristics which difference to initial state remained higher after 30 years.
This confirms conclusions drawn for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay between functional and taxonomic diversities meant the rapid recovery of dominant species while this of rare ones would be much longer.
Besides, it meant that species missing to plots taxonomic recovery would be functionally redundant with those already in place.
The functional overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006], is then not trivially recovered after disturbance and as a determinant of forests resilience this questions forests resilience to multiple disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2005].

>>>>

Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)).

@@

For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig6)). 
Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.

@@

The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2012).
Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).

The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth with a sharp increase until 15 years after disturbance.
Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
Communities composition after disturbance is either due to the old, pre-disturbance survivors or to the recruited trees.
The composition of old survivors proved to mirror the initial community [@Herault2018], the observed composition turnover would stem from an enhanced the growth and survival of previously infrequent species which, along with the IDH, reorganizes the typical high dominance structure of hyperdiverse mature forests.
This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
Post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].

@@

Plots functional diversity had recovered initial values for almost all plots 30 years after disturbance. 
It followed a hump-back trajectory that was mirror by all leaf, stem and life history traits, to the exception of leaf chlorophyll content.
This recovery of communities functional characteristics would go along the recovery of ecosystems processes [@Guariguata2001], as functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005].
Similar hump-back trajectories were followed by communities taxonomic and functional composition which argued for the convergence of communities composition maintaining the initial difference after the recovery cycle [@Hubbell1999;@Molino2001;@Baraloto2012a].
Besides, desipte a homogeneous trend among plots, the trajectories proved quite variable among plots, specifically after intense disturbance.
This meant a strong dependency of plots recovery time and trajectory to initial composition or abiotic conditions.

#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.

@@

Composition trajectories, though, translated marked differences in communities composition as the functional diversity had already recovered.
Such delay between functional and taxonomic dynamics, already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay meant a faster recovery of dominant species while this of infrequent ones would be much longer.
These infrequent species missing to plots taxonomic recovery would consequently be functionally redundant with those already in place.
The functional overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006], is then not trivially recovered after disturbance and as a determinant of forests resilience this questions forests resilience to multiple disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
